---
title: "Homework 1"
author: "Tinglei Wu"
date: "1/25/2022"
output: html_document
---
## load prostate data
```{r}
prostate <- 
  read.table(url(
    'https://web.stanford.edu/~hastie/ElemStatLearn/datasets/prostate.data'))
```

## subset to training examples
```{r}
prostate_train <- subset(prostate, train==TRUE)
```

## plot lcavol vs lpsa
```{r}
plot_psa_data <- function(dat=prostate_train) {
  plot(dat$lpsa, dat$lcavol,
       xlab="log Prostate Screening Antigen (psa)",
       ylab="log Cancer Volume (lcavol)",
       pch = 20)
}
plot_psa_data()
```

############################
## regular linear regression
############################

## loss function for l2
```{r}


L2_loss <- function(y, yhat)
  (y-yhat)^2

```

## fit simple linear model using numerical optimization
```{r}
fit_lin <- function(y, x, loss=L2_loss, beta_init = c(-0.51, 0.75)) {
  err <- function(beta)
    mean(loss(y,  beta[1] + beta[2]*x))
  beta <- optim(par = beta_init, fn = err)
  return(beta)
}
```

## make predictions from linear model
```{r}
predict_lin <- function(x, beta)
  beta[1] + beta[2]*x
```

## fit linear model
```{r}
lin_beta <- fit_lin(y=prostate_train$lcavol,
                    x=prostate_train$lpsa,
                    loss=L2_loss)
```

## compute predictions for a grid of inputs
```{r}
x_grid <- seq(min(prostate_train$lpsa),
              max(prostate_train$lpsa),
              length.out=100)
lin_pred <- predict_lin(x=x_grid, beta=lin_beta$par)
```

## plot data
```{r}
plot_psa_data()
## plot predictions
lines(x=x_grid, y=lin_pred, col='darkgreen', lwd=2)

## do the same thing with 'lm'

lin_fit_lm <- lm(lcavol ~ lpsa, data=prostate_train)


## make predictins using 'lm' object

lin_pred_lm <- predict(lin_fit_lm, data.frame(lpsa=x_grid))



## plot predictions from 'lm'

lines(x=x_grid, y=lin_pred_lm, col='pink', lty=2, lwd=2)

```

##################################
## try modifying the loss function
##################################

## custom loss function
```{r}
custom_loss <- function(y, yhat)
  (y-yhat)^2 + abs(y-yhat)
```

## plot custom loss function
```{r}
err_grd <- seq(-1,1,length.out=200)
plot(err_grd, custom_loss(err_grd,0), type='l',
     xlab='y-yhat', ylab='custom loss')
```

## fit linear model with custom loss
```{r}
lin_beta_custom <- fit_lin(y=prostate_train$lcavol,
                    x=prostate_train$lpsa,
                    loss=custom_loss)

lin_pred_custom <- predict_lin(x=x_grid, beta=lin_beta_custom$par)
```

## plot data
```{r}
plot_psa_data()
## plot predictions from L2 loss
lines(x=x_grid, y=lin_pred, col='darkgreen', lwd=2)
## plot predictions from custom loss
lines(x=x_grid, y=lin_pred_custom, col='pink', lwd=2, lty=2)
```


## Question 1:

```{r}
L1_loss <- function(y, yhat)
  abs(y-yhat)

L1_tilted <- function(y, yhat, tau)
  ifelse((y-yhat) > 0, tau*(y-yhat), (tau-1)*(y-yhat))

err_grd <- seq(-1,1,length.out=200)
plot(err_grd, L1_loss(err_grd,0), type='l',
     xlab='y-yhat', ylab='L1_loss', col = "red")

lines(err_grd, L2_loss(err_grd,0), type = 'l',
      xlab='y-yhat', ylab='L2_loss', col = "blue ")

lines(err_grd, L1_tilted(err_grd,0, 0.25), type = 'l',
      xlab='y-yhat', ylab='L1_tilted', col = "yellow")

lines(err_grd, L1_tilted(err_grd,0, 0.75), type = 'l',
      xlab='y-yhat', ylab='L1_tilted', col = "navy")

legend("topleft", legend=c("L1_loss", "L2_loss", "L1_tilted(0.25)", "L1_tilted(0.75)"),
       col=c("red", "blue", "yellow", "navy"), lty=1)
```


## Question 2:

```{r}
#L2
fit_lin_l2 <- function(y, x, loss=L2_loss, beta_init = c(-0.51, 0.75)) {
  err <- function(beta)
    mean(loss(y,  beta[1] + beta[2]*x))
  beta <- optim(par = beta_init, fn = err)
  return(beta)
}
predict_lin <- function(x, beta)
  beta[1] + beta[2]*x

lin_beta_l2 <- fit_lin_l2(y=prostate_train$lcavol,
                    x=prostate_train$lpsa,
                    loss=L2_loss)

x_grid <- seq(min(prostate_train$lpsa),
              max(prostate_train$lpsa),
              length.out=100)
lin_pred_l2 <- predict_lin(x=x_grid, beta=lin_beta_l2$par)

#L1
fit_lin_l1 <- function(y, x, loss=L1_loss, beta_init = c(-0.51, 0.75)) {
  err <- function(beta)
    mean(loss(y,  beta[1] + beta[2]*x))
  beta <- optim(par = beta_init, fn = err)
  return(beta)
}
predict_lin <- function(x, beta)
  beta[1] + beta[2]*x

lin_beta_l1 <- fit_lin_l1(y=prostate_train$lcavol,
                    x=prostate_train$lpsa,
                    loss=L1_loss)

x_grid <- seq(min(prostate_train$lpsa),
              max(prostate_train$lpsa),
              length.out=100)
lin_pred_l1 <- predict_lin(x=x_grid, beta=lin_beta_l1$par)

# L1_tilted(tau = 0.25)
fit_lin_tilted_le <- function(y, x, loss=L1_tilted, beta_init = c(-0.51, 0.75)) {
  err <- function(beta)
    mean(loss(y,  beta[1] + beta[2]*x, 0.25))
  beta <- optim(par = beta_init, fn = err)
  return(beta)
}
predict_lin <- function(x, beta)
  beta[1] + beta[2]*x

lin_beta_l1_tilted <- fit_lin_tilted_le(y=prostate_train$lcavol,
                    x=prostate_train$lpsa,
                    loss=L1_tilted)

x_grid <- seq(min(prostate_train$lpsa),
              max(prostate_train$lpsa),
              length.out=100)
lin_pred_l1_tilted <- predict_lin(x=x_grid, beta=lin_beta_l1_tilted$par)

# L1_tilted(tau = 0.75)
fit_lin_tilted_mo <- function(y, x, loss=L1_tilted, beta_init = c(-0.51, 0.75)) {
  err <- function(beta)
    mean(loss(y,  beta[1] + beta[2]*x, 0.75))
  beta <- optim(par = beta_init, fn = err)
  return(beta)
}
predict_lin <- function(x, beta)
  beta[1] + beta[2]*x

lin_beta_l1_tilted_mo <- fit_lin_tilted_mo(y=prostate_train$lcavol,
                    x=prostate_train$lpsa,
                    loss=L1_tilted)

x_grid <- seq(min(prostate_train$lpsa),
              max(prostate_train$lpsa),
              length.out=100)
lin_pred_l1_tilted_mo <- predict_lin(x=x_grid, beta=lin_beta_l1_tilted_mo$par)


plot_psa_data <- function(dat=prostate_train) {
  plot(dat$lpsa, dat$lcavol,
       xlab="log Prostate Screening Antigen (psa)",
       ylab="log Cancer Volume (lcavol)",
       pch = 20)
}
plot_psa_data()
lines(x_grid, lin_pred_l2, type = 'l', col = "red")

lines(x_grid, lin_pred_l1, type = 'l', col = "blue"
      )
lines(x_grid, lin_pred_l1_tilted, type = 'l', col = "yellow"
      )
lines(x_grid, lin_pred_l1_tilted_mo, type = 'l', col = "navy"
      )

legend("topleft", legend=c("lin_pred_l2", "lin_pred_l1", "lin_pred_l1_tilted(0.25)", "lin_pred_l1_tilted(0.75)"),
       col=c("red", "blue", "yellow", "navy"), lty=1)
```

##Question 3:

```{r}
#Nonlinear l2
fit_lin_l2_non <- function(y, x, loss=L2_loss, beta_init = c(-1.0, 0.0, -0.3)) {
  err <- function(beta)
    mean(loss(y,  beta[1] + beta[2]*exp(-beta[3]*x)))
  beta <- optim(par = beta_init, fn = err)
  return(beta)
}
predict_lin <- function(x, beta)
  beta[1] + beta[2]*exp(-beta[3]*x)

lin_beta_l2_non <- fit_lin_l2_non(y=prostate_train$lcavol,
                    x=prostate_train$lpsa,
                    loss=L2_loss)

x_grid <- seq(min(prostate_train$lpsa),
              max(prostate_train$lpsa),
              length.out=100)
lin_pred_l2_non <- predict_lin(x=x_grid, beta=lin_beta_l2_non$par)

#Nonlinear l1
fit_lin_l1_non <- function(y, x, loss=L1_loss, beta_init = c(-1.0, 0.0, -0.3)) {
  err <- function(beta)
    mean(loss(y,  beta[2]*exp(-beta[3]*x)))
  beta <- optim(par = beta_init, fn = err)
  return(beta)
}
predict_lin <- function(x, beta)
  beta[2]*exp(-beta[3]*x)

lin_beta_l1_non <- fit_lin_l1_non(y=prostate_train$lcavol,
                    x=prostate_train$lpsa,
                    loss=L1_loss)

x_grid <- seq(min(prostate_train$lpsa),
              max(prostate_train$lpsa),
              length.out=100)
lin_pred_l1_non <- predict_lin(x=x_grid, beta=lin_beta_l1_non$par)

#Nonlinear l1_tilted(tau = 0.25)
fit_lin_tilted_le_non <- function(y, x, loss=L1_tilted, beta_init = c(-1.0, 0.0, -0.3)) {
  err <- function(beta)
    mean(loss(y,  beta[1] + beta[2]*exp(-beta[3]*x), 0.25))
  beta <- optim(par = beta_init, fn = err)
  return(beta)
}
predict_lin <- function(x, beta)
  beta[1] + beta[2]*exp(-beta[3]*x)

lin_beta_l1_tilted_non <- fit_lin_tilted_le_non(y=prostate_train$lcavol,
                    x=prostate_train$lpsa,
                    loss=L1_tilted)

x_grid <- seq(min(prostate_train$lpsa),
              max(prostate_train$lpsa),
              length.out=100)
lin_pred_l1_tilted_non <- predict_lin(x=x_grid, beta=lin_beta_l1_tilted_non$par)

#Nonlinear l1_tilted(tau = 0.75)
fit_lin_tilted_mo_non <- function(y, x, loss=L1_tilted, beta_init = c(-1.0, 0.0, -0.3)) {
  err <- function(beta)
    mean(loss(y,  beta[1] + beta[2]*exp(-beta[3]*x), 0.75))
  beta <- optim(par = beta_init, fn = err)
  return(beta)
}
predict_lin <- function(x, beta)
  beta[1] + beta[2]*exp(-beta[3]*x)

lin_beta_l1_tilted_mo_non <- fit_lin_tilted_mo_non(y=prostate_train$lcavol,
                    x=prostate_train$lpsa,
                    loss=L1_tilted)

x_grid <- seq(min(prostate_train$lpsa),
              max(prostate_train$lpsa),
              length.out=100)
lin_pred_l1_tilted_mo_non <- predict_lin(x=x_grid, beta=lin_beta_l1_tilted_mo_non$par)
```

##Question 4:

```{r}

plot_psa_data <- function(dat=prostate_train) {
  plot(dat$lpsa, dat$lcavol,
       xlab="log Prostate Screening Antigen (psa)",
       ylab="log Cancer Volume (lcavol)",
       pch = 20)
}
plot_psa_data()
lines(x_grid, lin_pred_l2_non, type = 'l', col = "red")

lines(x_grid, lin_pred_l1_non, type = 'l', col = "blue"
      )
lines(x_grid, lin_pred_l1_tilted_non, type = 'l', col = "yellow"
      )
lines(x_grid, lin_pred_l1_tilted_mo_non, type = 'l', col = "navy"
      )

legend("topleft", legend=c("lin_pred_l2_non", "lin_pred_l1_non", "lin_pred_l1_tilted(0.25)_non", "lin_pred_l1_tilted(0.75)_non"),
       col=c("red", "blue", "yellow", "navy"), lty=1)


```