---
title: "hw4"
author: "Tinglei Wu"
date: "2/21/2022"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exercise 4:

## Part a:

* Since we know that if we want to predict the response for a test observation with X=0.6, we will use observations in the range [0.55,0.65], in this case, if x is between 0 and 1, then the observation we want to use are in the interval [x−0.05,x+0.05] which represents a length of 0.1 and a fraction of 10%. However, we would also consider a situation that if x is less than 0.05, which the observation interval becomes [0, x+0.05], because the interval cannot be negative. Another situation is that when x is greater than 0.95, so the interval would become [x-0.05, 1]. Therefore, the average fraction we will use to make the prediction is: 

\[ \int_{0.05}^{0.95} 10 \,dx  + \int_{0}^{0.05} 100x+5 \,dx + \int_{0.95}^{1} 105-100x \,dx = 9 + 0.375 + 0.375 = 9.75\]

Therefore, the average fraction of observations we would use for prediction is 9.75%


## Part b:

* When it becomes 2 features with p = 2, we can simply calculate the fraction of observations that we would use for prediction by using \[9.75\%^2 = 0.950625\].

## Part c:

* When the features become 100 with p = 100, it is the same thing for us to calculate the fraction of observations that we would use for prediction except the power would become 100: \[9.75\%^{100}\approx 0 \].



## Part d:

* As we can see from the previous questions, as the number of features increases, the fraction of observations that we would use for prediction decreases. When p becomes infinity, the fraction of observations that we would use for prediction becomes 0.

# Part e:

* Since it contains 10% of the trainning observations, when p = 1, length of each side of the hypercube is 0.1. When p = 2, the length of the each side of the hypercube is \[0.1^{1/2}\], when p = 100, the length of the each side of the hypercube is \[0.1^{1/100}\].


# Exercise 10:

## Part a:
```{r}
library(ISLR)
summary(Weekly)
cor(Weekly[, -9])
attach(Weekly)
plot(Volume)
```

* The Year and Volume variables seem to have very high positive correlation between each other,0.84194162, and the graph of Volume is also increasing over time.

## Part b:

```{r}
head(Weekly)

fit.glm <- glm(Direction ~ Lag1 + Lag2 + Lag3 + Lag4 + Lag5 + Volume, data = Weekly, family = binomial)
summary(fit.glm)
```

* From the results above, we can see that Lag2 is the only predictor that has a p-value lower than 0.05, so Lag2 is statistically significant.

## Part c:

```{r}
probs <- predict(fit.glm, type = "response")
pred.glm <- rep("Down", length(probs))
pred.glm[probs > 0.5] <- "Up"
table(pred.glm, Direction)


```


* Overall, the accuracy of the prediction is about (54+557)/1089 = 56.1%, thus the error rate of the prediction is about 43.9%.

## Part d:

```{r}
train <- (Year < 2009)
Weekly.20092010 <- Weekly[!train, ]
Direction.20092010 <- Direction[!train]
fit.glm2 <- glm(Direction ~ Lag2, data = Weekly, family = binomial, subset = train)
summary(fit.glm2)

probs2 <- predict(fit.glm2, Weekly.20092010, type = "response")
pred.glm2 <- rep("Down", length(probs2))
pred.glm2[probs2 > 0.5] <- "Up"
table(pred.glm2, Direction.20092010)
```

* In this case, we only use Lag2 as the predictor to predict the Direction, and the accuracy of the prediction is (9+56)/104 = 62.5%, thus the error rate of the prediction is 37.5%.

## Part e:

```{r}

library(MASS)
fit.lda <- lda(Direction ~ Lag2, data = Weekly, subset = train)
fit.lda
pred.lda <- predict(fit.lda, Weekly.20092010)
table(pred.lda$class, Direction.20092010)

```


* Using the LDA actually gives us the same result as glm, the accuracy of the prediction is (9+56)/104 = 62.5%, thus the error rate of the prediction is 37.5%. 

## Part f:

```{r}
fit.qda <- qda(Direction ~ Lag2, data = Weekly, subset = train)
fit.qda

pred.qda <- predict(fit.qda, Weekly.20092010)
table(pred.qda$class, Direction.20092010)
```

* Using the QDA gives us the accuracy of the prediction to be 61/104 = 58.65%, and the error rate of prediction is 41.35%. However, we can see that the model is only choosing Up as the answer and not even have one Down answer.

## Part g:

```{r}
library(class)
train.X <- as.matrix(Lag2[train])
test.X <- as.matrix(Lag2[!train])
train.Direction <- Direction[train]
set.seed(1)
pred.knn <- knn(train.X, test.X, train.Direction, k = 1)
table(pred.knn, Direction.20092010)





```


* The accuracy of prediction using KNN with k = 1 is (21+31)/104 = 50%, and thus the error rate of the prediction is also 50%. 

## Part h:

* From the previous results, we can see that the logistic regression and LDA have the best performances in terms of accuracy of the prediction. 


## Part i:


```{r}
# Logistic regression with Lag2:Lag4
fit.glm3 <- glm(Direction ~ Lag2:Lag4, data = Weekly, family = binomial, subset = train)
probs3 <- predict(fit.glm3, Weekly.20092010, type = "response")
pred.glm3 <- rep("Down", length(probs3))
pred.glm3[probs3 > 0.5] = "Up"
table(pred.glm3, Direction.20092010)

mean(pred.glm3 == Direction.20092010)

```

```{r}
# LDA with Lag2 interaction with Lag3
fit.lda2 <- lda(Direction ~ Lag3:Lag1, data = Weekly, subset = train)
pred.lda2 <- predict(fit.lda2, Weekly.20092010)
mean(pred.lda2$class == Direction.20092010)
```

```{r}
# QDA with Volume
fit.qda2 <- qda(Direction ~ Lag2 + Volume, data = Weekly, subset = train)
pred.qda2 <- predict(fit.qda2, Weekly.20092010)
table(pred.qda2$class, Direction.20092010)
mean(pred.qda2$class == Direction.20092010)
```


```{r}
# KNN k = 19
pred.knn2 <- knn(train.X, test.X, train.Direction, k = 19)
table(pred.knn2, Direction.20092010)
mean(pred.knn2 == Direction.20092010)


```

* After examine the combinations of predictors, the original logistic regression and LDA still have the best performaces in terms of accuracy of the prediction overall.






